Modern type-safe embedding of attribute grammars

### 2018-11-21 doc/Paper.tex
Update PDF to df9deb47d64763e52c95c3526f511282576a8198

\begin{abstract}

  % Introduction. In one sentence, what’s the topic?

@@

  In fact, maintaining a traditional attribute grammar system is such a hard
  effort that most such system that were proposed in the past are no longer
  active.

@@

  An increasingly popular alternative approach to the use of AGs relies on
  embedding them as first class citizens of general purpose programming
  languages~\cite{Oege00,DBLP:conf/sblp/MartinsFS13,erlangAGs,kiama,doaitse09icfp,balestrieri}.
  This avoids the burden of implementing a totally new language and associated
  system by hosting it in state-of-the-art programming languages. Following this
  approach one then exploits the modern constructions and infrastructure that
  are already provided by those languages and focus on the particularities of
  the domain specific language being developed.

  Functional zipper\cite{huet1997zipper} is a powerful abstraction which greatly
  simplifies the implementation of traversal algorithms performing a lot of
  local updates. Functional zippers have successfully been applied to constuct
  an attribute grammar embeddings in Haskell\cite{...}. Despite its elegance,
  this solution had a major drawback which prevented its use in real-world
  applications: attributes were not cached, but rather repeatedly recomputed
  which severely hurt performance. Recently, this flaw has been
  eliminated\cite{...} and replaced with a different one: the approach became
  intrusive, i.e. to benefit from the embedding user-defined data structures
  have to be adjusted.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Functional Zippers}
  Zipper is a data structure commonly used in functional programming for
  traversal with fast local updates. The zipper data structure was originally
  conceived by Huet\cite{huet1997zipper} in the context of trees. We will,
  however, first consider a simpler problem: a bidirectional list traversal.

@@

>>>>

\begin{abstract}
  % Introduction. In one sentence, what’s the topic?

@@

  In fact, maintaining a traditional attribute grammar system is such a hard
  effort that most such systems that were proposed in the past are no longer
  active.

@@

  An increasingly popular alternative approach to the use of AGs relies on
  embedding them as first class citizens of general purpose programming
  languages~\cite{Oege00,DBLP:conf/sblp/MartinsFS13,erlangAGs,kiama,doaitse09icfp,balestrieri}.
  This avoids the burden of implementing a totally new language and associated
  system by hosting it in state-of-the-art programming languages. Following this
  approach one then exploits the modern constructions and infrastructure that
  are already provided by those languages and focuses on the particularities of
  the domain specific language being developed.

  Functional zipper~\cite{huet1997zipper} is a powerful abstraction which greatly
  simplifies the implementation of traversal algorithms performing a lot of
  local updates. Functional zippers have successfully been applied to constuct
  an attribute grammar embedding in Haskell~\cite{DBLP:conf/sblp/MartinsFS13,MARTINS20162}. Despite its elegance,
  this solution had a major drawback which prevented its use in real-world
  applications: attributes were not cached, but rather repeatedly recomputed
  which severely hurt performance. Recently, this flaw has been
  eliminated~\cite{FERNANDES2018} and replaced with a different one: the approach became
  intrusive, i.e. to benefit from the embedding user-defined data structures
  have to be adjusted.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Functional Zippers}
  Zipper is a data structure commonly used in functional programming for
  traversal with fast local updates. The zipper data structure was originally
  conceived by Huet\cite{huet1997zipper} in the context of trees. We will,
  however, first consider a simpler problem: a bidirectional list traversal.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
